---
title: "Association rules"
author: "Akhilesh"
date: "August 18, 2018"
output: 
  md_document:
  variant: markdown_github
---

Association rule mining is a very interesting and important topic in retail analytics.Being able to find the associations between various products or services is the first step towards providing personalized recommendations.  
Apriori is one of the famous algorithms that is generally used for creating association rules and is also used for creating association rules for this problem. In this problem, i have tried my hand at finding the association rules between the different grocery items in the cart and visualizing the result in a user-friendly way.  

**Dataset: Grocery items**

Steps followed to create the association rules  

1.Data pre processing  
2.Running the apriori algorithm  
3.Visualization of the Output  

Let's create the rules that will help in identifying the associations     

```{r,echo = TRUE,message = FALSE,warning = FALSE}
library('R.utils')
library(arules)
library(arulesViz)
library(grid)
## File
file <- "groceries.txt"

temp = read.delim2(file, header = FALSE, sep = "\t", dec = ",")
n = dim(temp)[1]

temp$lists = strsplit(as.character(temp$V1),",")
temp <- tibble::rowid_to_column(temp, "ID")

df = data.frame(ID=integer(),
                items=character())

for (i in 1:n) {
k = data.frame(ID=i, y=temp[i,3])
names(k) = c('ID','items')
df = rbind(df,k)
}

# Turn ID into a factor
df$user = factor(df$ID)
items = split(x=df$items, f=df$ID)

## Remove duplicates ("de-dupe")
items = lapply(items, unique)

## Cast this variable as a special arules "transactions" class.
itemtrans = as(items,"transactions")

# Now run the 'apriori' algorithm
# Look at rules with support > .005 & confidence >.1 & length (# artists) <= 5
itemrules = apriori(itemtrans, 
                     parameter=list(support=.001, confidence=.4, maxlen=10))

inspect(subset(itemrules, subset=lift > 10 & confidence > 0.5))

sub1 = subset(itemrules, subset=confidence > 0.4 & support > 0.001)

plot(head(sub1, 10, by='lift'), method='graph')

```
Nodes with arrow pointing outwards denotes the items on LHS and nodes with arrow pointing inwards denotes the items on RHS  

**Gephi is a software that is very useful in creating graphs that are interpretable when it comes to association rules** [Gephi link](https://gephi.org/)

Following are some of the rules that are created using Gephi software  

```{r,fig.width=10, fig.height=20, echo = TRUE,message = FALSE, out.width="50%"}

library(knitr)
library(png)

myimages<-list.files("association_files", pattern = ".png", full.names = TRUE)
include_graphics(myimages)
```


**Association rules**  

Following are the rules that came out to be siginificant based on the support and lift that we considered:  

1.There is a high probability for people to red/blush wine if they buy bottled beer and liquor  
2.Buying soda and popcorn has a high association with salty snack  
3.Processed cheese and white bread has a high association with buying ham based on the lift   
4.Some rules are obvious like buying flour and baking powder has a high association with sugar  
5.Dairy products are associated with each other    
6.Ham, processed cheese and butter are highly associated  

By using Gephi,we can create a dynamic interactive viusalization that can incorporate a huge number of rules and would help in deriving insights for a particular category or department.  

Hope you enjoyed going through the analysis!

